The \frb\ Observing Campaign: Multi-Telescope Radio Burst and Implications for the FRB Population

### 2016-12-19 repeater-vladata.tex
draft circulated to group. has conclusions. fixed latex formatting issues.

Observations in August and Septmeber were searched by a prototype version of \rf\footnote{See \url{http://realfast.io}.}. \rf\ is a real-time, fast imaging transient search system. The current, prototype runs on existing, CPU-based hardware of the VLA correlator backend, while the future \rf\ will run on a dedicated GPU cluster. The transient search pipeline software is called \emph{rtpipe}\footnote{See \url{https://github.com/caseyjlaw/rtpipe}} and is mostly written in Python (detail in \citet{2015ApJ...807...16L}). Images were formed for each integration at DMs of 0, 546, 556.9, 560, and 565 pc cm$^{-3}$. Gain calibration is read from the "telcal" system, which uses phase-only calibration on the previous gain calibrator. A flux scale is calculated for each spectral window from an observation on {\color{red} xx (Bryan)} and applied to all burst spectra.

Burst detections and localizations were made within hours of data being recorded. The transient search starts when data are recorded, but this prototype of \rf\ is a factor of several times slower than real-time, so we refer to the detection as "quasi real-time". Each image with a pixel higher than 6.4$\sigma$ is saved in a summary form as a check of data quality. For each image with a pixel higher than 7.4$\sigma$, \rf\ generates a more detailed candidate visualization with an image and spectrum. More detailed analysis, including improved calibration and localization, is conducted offline.

@@

{\color{red} copied from the first paper}
During the joint Arecibo-VLA campaign, Arecibo observed with the L-wide receiver, which has an observational frequency range of 1.15 to 1.73~GHz and a full width at half maximum beam size of 3.3 arcmin. The PUPPI pulsar backend was used to record total intensity spectra with time and frequency resolutions of 10.24~$\mu$s and 1.5625~MHz, respectively, and full Stokes polarization information. Each frequency channel was coherently dedispersed to 557 pc cm$^{-3}$, thereby eliminating intra-channel dispersion smearing. PUPPI covers a total of 800~MHz of bandwidth centred at 1380.78125~MHz, but only $\sim$ 620~MHz of this band is usable due to radio frequency interference and receiver sensitivity roll-off at the band edges.

@@

We observed \frb with the Arcminute MicroKelvin Imager Large Array (AMI-LA; Zwart et al. 2008) for 3 hours each on four epochs starting at MJDs 57643.3351, 57645.3317, 57648.3352, and 57649.3353. Observations were made with the new digital correlator having 4096 channels across a 5 GHz bandwidth between 13--18 GHz with a 1s dump time. The phase calibrator, J0518+3306, was observed every 12 minutes for about 1.5 minutes. The AMI-LA data were binned to eight 0.625 GHz channels and processed (RFI excision and calibration) with a fully-automated pipeline, AMI-REDUCE \citep[e.g.,][]{2013MNRAS.429.3330P}. Daily measurements of 3C48 and 3C286 were used for the absolute flux calibration, which is good to about 10\%.

@@

All but two of the best-fit Gaussians are centered inside the 3~GHz band and most appear contained by the 1~GHz wide band. This is consistent with previous detections of \frb by Arecibo, which showed quasi-broadband structure \citep{2016arXiv160308880S} and large variation in the implied spectral index \citep{2014ApJ...790..101S}. A detailed discussion of high-resolution \frb\ burst spectra seen by the Arecibo Observatory will be presented elsewhere \citep{WEIRD}.

@@

 **Preliminary** Figure \ref{fig:dmmodel} compares the 95\% confidence intervals on the DM for all nine VLA bursts. The DM confidence intervals {\color{red} are/are not} consistent with a single DM... %, which suggests that there are apparent DM changes between bursts. This change could be caused by intrinsic burst structure or actual changes to the DM column density.

High-resolution dynamic spectra at 1.4~GHz have discovered spectro-temporal structure that can bias the apparent DM differently from burst to burst \citep{WEIRD}. Observations at 1.4~GHz have measured DM in the range from xx to yy pc cm$^{-3}$, which is **smaller than/consistent with** the range of observed values seen by the VLA at 3~GHz. The FRB emission process has not yet been defined, so it is not clear how the intrinsic structure can bias DM as a function of frequency.

\begin{figure}[htb]
\begin{center}
\includegraphics[width=0.9\columnwidth]{dmmodel}
\caption{**Preliminary** Comparison of 95\% ($1\sigma$) confidence intervals for DM for nine bursts detected by the VLA.
\label{fig:dmmodel}}
\end{center}
\end{figure}

@@

\subsubsection{Spectral Autocorrelation}
\label{sec:auto}
Autocorrelation of the burst signal (both temporal and spectral) can be used to infer both intrinsic properties and modulation due to scintillation \citep{CORDES}. While VLA spectral resolution of 4~MHz is relatively coarse, the Galactic electron density model predicts a diffractive scintillation bandwidth of {\color{red} xx MHz at 3~GHz (Jim?)} \citep{2002astro.ph..7156C}. Figure \ref{fig:acf} shows the spectral autocorrellation for the strongest burst (MJD 57633.68). No excess correlation is found near the channel resolution, which limits the correlation width to less than 4~MHz.

\begin{figure}[htb]
\begin{center}
\includegraphics[width=0.9\columnwidth]{acf_57633_scan7}
\caption{The spectral autocorrelation for the burst from \frb\ at MJD 57633.68. Blue shows the autocorrelation for the burst and green shows an autocorrelation of a representative spectrum with no burst.
\label{fig:acf}}
\end{center}
\end{figure}

\subsubsection{Circular Polarization}
The VLA S-band recievers natively measure circular polarization, although for observing efficiency we chose not include polarization calibration procedures. Crude constraints on circular polarization are possible by comparing the burst intensity in right and left-hand polarized data products. The apparent circular polarization fraction ($(RR-LL)/(RR+LL)$) for the most significant bursts are all less than 3\%. \frb was located 2.3 arcmin away from pointing center, where systematic effects have been measured as large as 3\% (Perley et al 2016, VLA memo). Given that systematic effects dominate the apparent circular polarization, we conclude that the true fractional circular polarization is less than 3\%.

\subsection{Temporal Statistics}
\label{sec:temp}
Burst detections were made very inhomogenously through the larger (63 hours) observing campaign of FRB 121102. In the first 30 hours of observing at S-band no bursts were detected, while nine bursts were detected in the last 27 hours of S-band observing. The data quality is high, so the inhomogeneous burst distribution shows that the burst detection probability was not stationary. Assuming that the burst detection probability follows a Poisson distribution, the nondetection in the first half of S-band limits the FRB rate to $\rm{R}<0.1$ hour$^{-1}$ (95\% confidence limit). The mean detection rate for the last part of the campaign was $\rm{R}=0.3$ hour$^{-1}$.

@@

This uniform sample of burst luminosities can be used to infer more general properties of \frb. We treat the detection probability as a Poisson process with a rate that scales with the luminosity as a powerlaw. Figure \ref{fig:lumd} shows the probability distribution for the powerlaw amplitude and index for the sample of nine VLA bursts. The luminosity powerlaw index $1\sigma$ bounds range from --0.5 to +0.4; for a cumulative luminosity distribution, this corresponds to a powerlaw index of roughly --0.6. The flatness of the distribution is also clear from range of significance of the nine bursts (12 to 179$\sigma$). All bursts are much brighter than the sensitivity limit, so the distribution is not affected by a completeness limit.

\begin{figure}[htb]
\begin{center}
\includegraphics[width=0.9\columnwidth]{luminosity_disn}
\caption{Probability distribution of powerlaw luminosity model for the nine VLA bursts from \frb. Powerlaw referenced to $L=2\times10^{38}$\ erg.
\label{fig:lumd}}
\end{center}
\end{figure}

@@

These simulations confirm previous work that shows that in many circumstances the cumulative flux distribution follows a Euclidean distribution \citep[index = --1.5;][]{2016MNRAS.462..941L}. To reproduce the flatness (index$\approx-1$) of the observed flux distribution, we need to incorporate other effects. 
**in progress** (new figure?)

@@

\subsection{Emission Physics and Burst Energetics}
The measurement of a redshift to the host of \frb allows us to measure the burst luminosity. We now know that \frb\ is detectable with the VLA out to $z=0.7$ and much farther with Arecibo. This confirms the results presented in \citet{LOC}, which quoted a luminosity on the order of $10^{38}$\ erg. The more refined analysis presented here measures somewhat higher burst fluxes, so the brightest VLA burst luminosity is now on the order of $10^{40}$\ erg.

@@

\subsection{FRB Rate}
% cut from gemini paper. TODO: incorporate referee comments on rate scale. Sarah?
Assuming that \frb is from the same population as the other FRBs, the measurement of a distance for the host of \frb allows us to re-evaluate the cosmic volume and event rate per galaxy for the FRB population. The estimated projected FRB rate $R_p$ understates the true rate by a beaming fraction $\Omega_b$ \citep[$\sim$10\%;][]{1998MNRAS.298..625T}.. For a comoving volume $V(z)$ and galaxy number density $\Phi(M)$, the rate per galaxy is $R_{FRB} = R_p /(\Omega_b \Phi(M)V (z))$.

The first such rate estimate was made by \citet{2013Sci...341...53T}, who used the measured DM to estimate a characteristic distance for their sample of 4 FRBs. They assumed that all of the extragalactic component of the DM was caused by the IGM and scaled as $DM\approx z\times1200 \rm{pc}\ \rm{cm}^{-3}$ \citep{2003ApJ...598L..79I,2004MNRAS.348..999I}. They also calculated the number of galaxies by assuming a characteristic $L_*$ galaxy \citep[corresponding to stellar mass $M_* \approx 1010.66 M$][]{2012MNRAS.421..621B}.

@@

\section{Conclusions}
Prior to being established as a cosmological source, the low Galactic latitude of \frb\ made it a somewhat compromised member of the FRB class. With its cosmological distance firmly established, \frb\ now serves as a new kind of standard by which FRBs are defined.

 % include timeline and science scope for new localizations? realfast plug and reference cosmology papers, contingency on DMhost, etc.

>>>>

Observations in August and Septmeber were searched by a prototype version of \rf\footnote{See \url{http://realfast.io}.}. \rf\ is a real-time, fast imaging transient search system. The current, prototype runs on existing, CPU-based hardware of the VLA correlator backend, while the future \rf\ will run on a dedicated GPU cluster. The transient search pipeline software is called \emph{rtpipe}\footnote{See \url{https://github.com/caseyjlaw/rtpipe}} and is mostly written in Python (detail in \citet{2015ApJ...807...16L}). Images were formed for each integration at DMs of 0, 546, 556.9, 560, and 565 pc cm$^{-3}$. Gain calibration is read from the "telcal" system, which uses phase-only calibration on the previous gain calibrator. A flux scale is calculated for each spectral window from an observation on {\color{red}xx (Bryan)} and applied to all burst spectra.

Burst detections and localizations were made within hours of data being recorded. The transient search starts when data are recorded, but this prototype of \rf\ is a factor of several times slower than real-time, so we refer to the detection as "quasi real-time". Each image with a pixel higher than 6.4$\sigma$\ is saved in a summary form as a check of data quality. For each image with a pixel higher than 7.4$\sigma$, \rf\ generates a more detailed candidate visualization with an image and spectrum. More detailed analysis, including improved calibration and localization, is conducted offline.

@@

{\color{red}Jason/Shami: ok if copied from the first paper?}
During the joint Arecibo-VLA campaign, Arecibo observed with the L-wide receiver, which has an observational frequency range of 1.15 to 1.73~GHz and a full width at half maximum beam size of 3.3 arcmin. The PUPPI pulsar backend was used to record total intensity spectra with time and frequency resolutions of 10.24~$\mu$s and 1.5625~MHz, respectively, and full Stokes polarization information. Each frequency channel was coherently dedispersed to 557 pc cm$^{-3}$, thereby eliminating intra-channel dispersion smearing. PUPPI covers a total of 800~MHz of bandwidth centred at 1380.78125~MHz, but only $\sim 620$\ MHz of this band is usable due to radio frequency interference and receiver sensitivity roll-off at the band edges.

@@

We observed \frb\ with the Arcminute MicroKelvin Imager Large Array (AMI-LA; Zwart et al. 2008) for 3 hours each on four epochs starting at MJDs 57643.3351, 57645.3317, 57648.3352, and 57649.3353. Observations were made with the new digital correlator having 4096 channels across a 5 GHz bandwidth between 13--18 GHz with a 1s dump time. The phase calibrator, J0518+3306, was observed every 12 minutes for about 1.5 minutes. The AMI-LA data were binned to eight 0.625 GHz channels and processed (RFI excision and calibration) with a fully-automated pipeline, AMI-REDUCE \citep[e.g.,][]{2013MNRAS.429.3330P}. Daily measurements of 3C48 and 3C286 were used for the absolute flux calibration, which is good to about 10\%.

@@

All but two of the best-fit Gaussians are centered inside the 3~GHz band and most appear contained by the 1~GHz wide band. This is consistent with previous detections of \frb\ by Arecibo, which showed quasi-broadband structure \citep{2016arXiv160308880S} and large variation in the implied spectral index \citep{2014ApJ...790..101S}. A detailed discussion of high-resolution \frb\ burst spectra seen by the Arecibo Observatory will be presented elsewhere \citep{WEIRD}.

@@

Figure \ref{fig:dmmodel} compares the 95\% confidence intervals on the DM for all nine VLA bursts to the range of DM previous observed \citep{2014ApJ...790..101S,2016arXiv160308880S}. The DM confidence intervals are not consistent with each other or a single DM. This suggests that there are apparent DM changes between bursts, either due to intrinsic burst structure or actual changes to the DM column density. Elsewhere, we present high-resolution dynamic spectra at 1.4~GHz that show how intrinsic spectral structure can bias the apparent DM of individual bursts \citep{WEIRD}. The FRB emission process has not yet been defined, so it is not clear how the intrinsic structure can bias DM as a function of frequency.

\begin{figure}[htb]
\begin{center}
\includegraphics[width=0.9\columnwidth]{dmmodel}
\caption{DM 95\% confidence intervals of the nine 3~GHz bursts detected by the VLA. The grey band shows the range of DM previously detected at 1.4~GHz ($\pm3$\ pc cm$^{-3}$).
\label{fig:dmmodel}}
\end{center}
\end{figure}

@@

\subsubsection{Spectral Autocorrelation}
\label{sec:auto}
Autocorrelation of the burst signal (both temporal and spectral) can be used to infer both intrinsic properties and modulation due to scintillation \citep{CORDES}. While VLA spectral resolution of 4~MHz is relatively coarse, the Galactic electron density model predicts a diffractive scintillation bandwidth of {\color{red}Jim: xx MHz at 3~GHz} \citep{2002astro.ph..7156C}. Figure \ref{fig:acf} shows the spectral autocorrellation for the strongest burst (MJD 57633.68). No excess correlation is found near the channel resolution, which limits the correlation width to less than 4~MHz.

\begin{figure}[htb]
\begin{center}
\includegraphics[width=0.9\columnwidth]{acf_57633_scan7}
\caption{ {\color{red} Jim: worth keeping?} The spectral autocorrelation for the burst from \frb\ at MJD 57633.68. Blue shows the autocorrelation for the burst and green shows an autocorrelation of a representative spectrum with no burst.
\label{fig:acf}}
\end{center}
\end{figure}

\subsubsection{Circular Polarization}
The VLA S-band recievers natively measure circular polarization, although for observing efficiency we chose not include polarization calibration procedures. Crude constraints on circular polarization are possible by comparing the burst intensity in right and left-hand polarized data products. The apparent circular polarization fraction ($(RR-LL)/(RR+LL)$) for the most significant bursts are all less than 3\%. \frb\ was located 2.3 arcmin away from pointing center, where systematic effects have been measured as large as 3\% (Perley et al 2016, VLA memo). Given that systematic effects dominate the apparent circular polarization, we conclude that the true fractional circular polarization is less than 3\%.

\subsection{Temporal Statistics}
\label{sec:temp}
Burst detections were made very inhomogenously through the larger (63 hours) observing campaign of FRB 121102. In the first 30 hours of observing at S-band no bursts were detected, while nine bursts were detected in the last 27 hours of S-band observing. The data quality is high, so the inhomogeneous burst distribution shows that the burst detection probability was not stationary. Assuming that the burst detection probability follows a Poisson distribution, the nondetection in the first half of S-band limits the FRB rate to $\rm{R}<0.1$\ hour$^{-1}$\ (95\% confidence limit). The mean detection rate for the last part of the campaign was $\rm{R}=0.3$\ hour$^{-1}$.

@@

This uniform sample of burst luminosities can be used to infer more general properties of \frb. We treat the detection probability as a Poisson process with a rate that scales with the luminosity as a powerlaw. Figure \ref{fig:lumd} shows the probability distribution for the powerlaw amplitude and index for the sample of nine VLA bursts. The luminosity powerlaw index $1\sigma$\ bounds range from --0.5 to +0.4; for a cumulative luminosity distribution, this corresponds to a powerlaw index of roughly --0.6. The flatness of the distribution is also clear from range of significance of the nine bursts (12 to 179$\sigma$). All bursts are much brighter than the sensitivity limit, so the distribution is not affected by a completeness limit.

\begin{figure}[htb]
\begin{center}
\includegraphics[width=0.9\columnwidth]{luminosity_disn}
\caption{Probability distribution of powerlaw luminosity model for the nine VLA bursts from \frb. The powerlaw is referenced to $L=2\times10^{38}$\ erg.
\label{fig:lumd}}
\end{center}
\end{figure}

@@

These simulations confirm previous work that shows that in many circumstances the cumulative flux distribution follows a Euclidean distribution \citep[index $=-1.5$;][]{2016MNRAS.462..941L}. To reproduce the flatness (index $\approx-1$) of the observed flux distribution, we need to incorporate other effects. 
**in progress** (new figure?)

@@

\subsection{Emission Physics and Burst Energetics}
The measurement of a redshift to the host of \frb\ allows us to measure the burst luminosity. We now know that \frb\ is detectable with the VLA out to $z=0.7$\ and much farther with Arecibo. This confirms the results presented in \citet{LOC}, which quoted a luminosity on the order of $10^{38}$\ erg. The more refined analysis presented here measures somewhat higher burst fluxes, so the brightest VLA burst luminosity is now on the order of $10^{40}$\ erg.

@@

\subsection{FRB Rate}
% cut from gemini paper. TODO: incorporate referee comments on rate scale. Sarah?
{\color{red} Sarah: can you update this as a per-SFR rate?}
Assuming that \frb\ is from the same population as the other FRBs, the measurement of a distance for the host of \frb\ allows us to re-evaluate the cosmic volume and event rate per galaxy for the FRB population. The estimated projected FRB rate $R_p$\ understates the true rate by a beaming fraction $\Omega_b$\ \citep[$\sim$10\%;][]{1998MNRAS.298..625T}.. For a comoving volume $V(z)$\ and galaxy number density $\Phi(M)$, the rate per galaxy is $R_{FRB} = R_p /(\Omega_b \Phi(M)V (z))$.

The first such rate estimate was made by \citet{2013Sci...341...53T}, who used the measured DM to estimate a characteristic distance for their sample of 4 FRBs. They assumed that all of the extragalactic component of the DM was caused by the IGM and scaled as $DM\approx z\times1200 \rm{pc}\ \rm{cm}^{-3}$\ \citep{2003ApJ...598L..79I,2004MNRAS.348..999I}. They also calculated the number of galaxies by assuming a characteristic $L_*$\ galaxy \citep[corresponding to stellar mass $M_* \approx 1010.66 M$;][]{2012MNRAS.421..621B}.

@@

\section{Conclusions}
Prior to being established as a cosmological source, the low Galactic latitude of \frb\ made it a somewhat compromised member of the FRB class. With its cosmological distance firmly established, \frb\ now serves as a new kind of standard by which FRBs are defined. That fact, combined with an abundance of data collected during its high activity state, allows us to think more generally about the physics of broader FRB population.

We presented the first multi-telescope detection (Arecibo and VLA) of a FRB. By detecting this burst from 1.2 to 4.5~GHz, we have demonstrated that some bursts have broad spectral structure. However, three other VLA bursts are undetected by Arecibo. That, combined with the measurement of spectra within the VLA band, suggest that the typical burst has a characteristic Gaussian spectral shape with width $\sim500$\ MHz.

**Flux/lum distribution**

Analysis of the VLA burst times shows that the burst rate is highly variable. This shows that past constraints on FRB repetition are weaker than previously inferred. The combination of relatively narrow spectral structure, flat luminosity and flux distributions, and variable burst rate suggests that repeated observations, wide bandwidth, and large instantaneous field of view all improve the chance of detection.

New, arcsecond-scale localizations will continue to be highly informative, given that FRB hosts may be faint and can confuse the inference of a properties of the intergalactic medium \citep{2014ApJ...783L..35D}. \frb\ was localized within hours by a prototype version of \rf, but an expanded \rf\ system is now under construction. This platform will search a TB/hour data stream in real time in parallel with ongoing VLA observations, potentially detecting and localizing multiple FRBs per year. {\color{red} Sarah: any refinement to this statement?}
